---
title: "Lab1"
author: "Connor Quiroz"
date: "2024-09-27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 1

> Downloaded R, completed lab in rmd file.

# Exercise 2

> Named file as Lab1_Quiroz.rmd and created Rproj file to set working directory.

# Excercise 3

```{r e3}
# LAB1: Intro R Lab in QSCI482
# September 27 2024 created by Connor Quiroz
```

# Exercise 4

```{r e4}
a = 0.00021
b = 3.1
L = 30
W <- signif(a * (L**b),1)
```

> Answer: 1 significant digit; 8kg

# Exercise 5

```{r e5}
# Create  and call Antarctic blue whales vector
bluelengths <- rnorm(100, mean = 84.4, sd = 4.07)
bluelengths

# Histogram of Antarctic blue whales lengths
hist(bluelengths)

# Max value of Antarctic blue whales vector
max(bluelengths)

# Trying 1000, 10000, and 100000 resamples from a normal distribution of Antarctic blue whales
bluelengths <- rnorm(1000, mean = 84.4, sd = 4.07)
max(bluelengths)
bluelengths <- rnorm(10000, mean = 84.4, sd = 4.07)
max(bluelengths)
bluelengths <- rnorm(100000, mean = 84.4, sd = 4.07)
max(bluelengths)
```

> Answer: maximum values of blue whales vector increases with higher sample sizes

# Exercise 6

### 6.1 Load in data
```{r e6.1}
cdata <- read.csv("CommuteTime.csv")
```

### 6.2 Print data

```{r e6.2}
print(cdata)
```

### 6.3 Extract first column

```{r e6.3}
commutetime <- cdata[1]
```

### 6.4 Mean and standard deviation

```{r e6.4}
Mean <- signif(mean(commutetime$TotalCommuteMin),2)
SD <- signif(sd(commutetime$TotalCommuteMin),2)
```

> Answer: Mean commute time for Trevor branch to work: `r Mean`min; standard deviation: `r SD`min.

### 6.5

```{r e6.5}
pnorm(60, mean = Mean, sd = SD)
```
> Answer: Probability that Trevor's commute time will NOT exceed 60 minutes and be early for work: 0.999.

### 6.6

```{r e6.6}
1-pnorm(60, mean = Mean, sd = SD)
```
> Answer: Probability that Trevor's commute time WILL exceed 60 minutes and be early for work: 0.0001.